---
title: "R-Basics"
output: html_document
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(comment = NA)
```


# What is `R`

As found on the [project's website](http://cran.us.r-project.org/):  


>  R is ‘GNU S’, a freely available language and environment for statistical computing and graphics which provides a wide variety of statistical and graphical techniques: linear and nonlinear modelling, statistical tests, time series analysis, classification, clustering, etc.  
  
  
I like to think of `R` as a 

>  Statistical programming language  

In that you can:

-  manipulate the file system  
-  interface with modern Web APIs  
-  read/write data to most contemporary database formats  
-  and oh by the way, you can automate statistical analyses and reports


## Basic functionality

At it's simplest form, `R` is just a calculator

```{r}
2+2
5 / 3
5 ** 2
5 ^ 2
```

While you may need to learn some basic syntax, the point from the code blocks above is that `R` is (most likely) capable of handling what you learned what you learned in math class.


##  A soft introduction into Object Oriented Programing

While typing `interactively` into a calculator (or command line prompt) can be useful, you will more than likely want to __save your work__ while thinking through a problem.  Instead of just asking for the sum of `2+2`, we can save the value for later!

```{r}
x = 2 +2 
x
```


From this point forward, we can always refer to the value of `x`.  In our simple example, `x`'s value is equal to `4`, but we could have set it to a phrase like `x = "my value is 4" ` or something more complicated.  We will cover this later.




